---
title: "Example Usage with Single Continuous Treatment Observational Data"
author:
- "Itamar Caspi"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: haddock
    theme: journal
    toc: yes
    toc_depth: 4
    toc_float: yes
abstract: |
  TBA
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(eval = TRUE,
                      echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

```


We applied our technique to Dominick’s dataset, a popular historical dataset of store-level orange juice prices and sales provided by University of Chicago Booth School of Business.

The dataset is comprised of a large number of covariates $W$, but researchers might only be interested in learning the elasticity of demand as a function of a few variables $x$ such as income or education.

We applied the DMLCateEstimator to estimate orange juice price elasticity as a function of income, and our results, unveil the natural phenomenon that lower income consumers are more price-sensitive.

## Packages

```{r}

library(reticulate)
use_virtualenv("myenv")

```


```{r}

library(tidymodels)
library(tidyverse)
library(here)

```


## Data

```{r}

# Read from file
oj_raw <- read_csv(here("data", "oj_large.csv"))

# Alternatively, read from the web
# oj_raw <- read_csv("https://msalicedatapublic.blob.core.windows.net/datasets/OrangeJuice/oj_large.csv")

head(oj_raw)

```


# Preprocessing

```{r}

oj_rec <- recipe(logmove ~ ., data = oj_raw) %>% 
  step_normalize(
    INCOME, AGE60, EDUC, ETHNIC, INCOME,
    HHLARGE, WORKWOM, HVAL150, SSTRDIST,
    SSTRVOL, CPDIST5, CPWVOL5
  ) %>% 
  step_dummy(brand, one_hot = TRUE) %>% 
  step_log(price) %>% 
  prep() %>% 
  juice()

```


```{r}

Y <- oj_rec %>%
  pull(logmove) %>%
  as.array() %>% 
  unname()
  
D <- oj_rec %>%
  pull(price) %>%
  as.array() %>% 
  unname()

X <- oj_rec %>%
  select(INCOME) %>%
  as.matrix() %>% 
  unname()

W <- oj_rec %>%
  select(AGE60:CPWVOL5, starts_with("brand")) %>%
  as.matrix() %>% 
  unname()
```

Generate testdata
```{r}

min_income <- -1
max_income <- 1
delta      <- (max_income - min_income) / 100

X_test <- seq(max_income, min_income, -delta) %>%
  as.matrix()

```


# Estimate effect

```{python}

from econml.dml import DMLCateEstimator 
from sklearn.ensemble import RandomForestRegressor,RandomForestClassifier

```


```{python}

est = DMLCateEstimator(model_y=RandomForestRegressor(),model_t=RandomForestRegressor())

est.fit(r.Y, r.D, r.X, r.W)

te_pred = est.effect(r.X_test)

```

Plot oranje Juice elasticity as a function of income
```{r}

cate_effect <- tibble("X_test" = X_test, "te_pred" = py$te_pred)

cate_effect %>% 
  ggplot(aes(X_test, te_pred)) +
  geom_line() +
  labs(x = "Scale(Income)",
       y = "Orange Juice Elasticity",
       title = "Orange Juice Elasticity vs Income")

```


## Bootstrap Confidence Intervals


```{python}

from econml.bootstrap import BootstrapEstimator

```


```{python}

boot_est = BootstrapEstimator(DMLCateEstimator(model_y=RandomForestRegressor(),model_t=RandomForestRegressor()),n_bootstrap_samples=20) 

boot_est.fit(r.Y, r.D, r.X, r.W)

te_pred_interval = boot_est.const_marginal_effect_interval(r.X_test, lower=1, upper=99)

```



```{r}

cate_effect_interval <- cate_effect %>% 
  mutate(
    te_pred_down = py$te_pred_interval[[1]],
    te_pred_up   = py$te_pred_interval[[2]]
  )

head(cate_effect_interval)
```


```{r}

cate_effect_interval %>% 
  ggplot(aes(X_test, te_pred)) +
  geom_ribbon(aes(ymin = te_pred_down, ymax = te_pred_up),
              fill = "grey70") +
  geom_line() +
  labs(x = "Scale(Income)",
       y = "Orange Juice Elasticity",
       title = "Orange Juice Elasticity vs Income",
       fill  = "1-99% CI")

```
